Ignition of Fuel Beds by Firebrands

### 2022-05-23 manuscript4.tex
Draft of results for the multi-heater study

    Further insight into the influence of heater configuration and wind speed on ignition is gained from the results of the computational efforts. Figure~\ref{fig:multiHeaterCFD} shows the averaged velocity streamlines and temperature distribution in the fluid flow around the heaters for each heater configuration. Air flows from left to right in Figure~\ref{fig:multiHeaterCFD}. The numbers inside the circular heater boundaries correspond to the assigned temperature of the heater during the calculation. The assigned surface temperatures correspond to the heater temperature anticipated to produce a 50\% probability of ignition for the wind speed and heater conditions. Heater boundaries with blue circles inside them correspond to cases where the heater was not heated and acted as an inert object. The arrows indicate the observed location of ignition, when observed. Ignition locations were not observed for every case, however, sufficient ignition locations were recorded to provide insight into controlling parameters.
        \begin{figure}
            \centering
            \includegraphics[width=\columnwidth]{Figures/quarter_circle.pdf}
            \caption{Time averaged velocity streamlines and temperature profiles for each configuration. The numbers inside of the heater boundaries identify surface temperature of the heater. Blue circles indicate that the heater was unheated and acting as an inert object. Arrows typical ignition locations.}
            \label{fig:multiHeaterCFD}
        \end{figure}

>>>>

    Further insight into the influence of heater configuration and wind speed on ignition is gained from the results of the computational efforts. Figure~\ref{fig:multiHeaterCFD} shows the averaged velocity streamlines and temperature distribution in the fluid flow around the heaters for each heater configuration. Air flows from left to right in Figure~\ref{fig:multiHeaterCFD}. The solid black lines separate the figure into quadrants with the same heater spacing and wind speed. Each quadrant contains two images, one where both heaters are actively heated and one where only the downstream heater is actively controlled. The numbers inside the circular heater boundaries correspond to the assigned temperature of the heater during the calculation. The assigned surface temperatures correspond to the heater temperature anticipated to produce a 50\% probability of ignition for the wind speed and heater conditions. Heater boundaries with blue circles inside them correspond to cases where the heater was not heated and acted as an inert object. The arrows indicate the observed location of ignition, when observed. Ignition locations were not observed for every case, however, sufficient ignition locations were recorded to provide insight into controlling parameters.
        \begin{figure}
            \centering
            \includegraphics[width=\columnwidth]{Figures/quarter_circle.pdf}
            \caption{Time averaged velocity streamlines and temperature profiles for each configuration. The numbers inside of the heater boundaries identify surface temperature of the heater. Blue circles indicate that the heater was unheated. Arrows show typical ignition locations.}
            \label{fig:multiHeaterCFD}
        \end{figure}
    The most significant difference in ignition probability from the experimental results was observed when the wind speed was increased from 0.5\si{\meter\per\second} to 5.8\si{\meter\per\second}. The computational results for the 0.5\si{\meter\per\second} cases are represented in to top two quadrants of Figure~\ref{fig:multiHeaterCFD} and the results for the 5.8\si{\meter\per\second} cases are represented in the bottom two quadrants of Figure~\ref{fig:multiHeaterCFD}. When comparing differences between the streamlines for the two different wind speeds the cases with a wind speed 5.8\si{\meter\per\second} have recirculation zones near the intersection of the fuel bed and the heater and the 0.5\si{\meter\per\second} cases do not. Figure~\ref{fig:streamlineComparison} shows an enlarged images of the streamlines for two cases from Figure~\ref{fig:multiHeaterCFD} where both heaters are active and spaced one diameter apart. Considering the regions on the upstream (left) side of the heaters the 5.8\si{\meter\per\second} case shown in Figure~\ref{subfig:comparisonHighWind} has distinct recirculation zones. Notably, ignition observed during experiments that correspond to Figure~\ref{subfig:comparisonHighWind} and Figure~\ref{fig:multiHeaterCFD} (bottom left quadrant, bottom plot) was observed to occur on the upstream side of the upstream heater which corresponds to the leftmost recirculation zone on Figure~\ref{subfig:comparisonHighWind}. This is perhaps unexpected as the leading edge of the downstream heater also has a recirculation zone near the fuel bed and a secondary recirculation zone exists between the heaters that creates a large region of fluid at temperatures above ambient which would facilitate ignition. Instead it appears that the conditions in the furthest upstream recirculation zone are more favorable. A similar location for ignition was observed for single heater configurations in Chapter~\ref{part:manuscript2}. The similar observations of ignition location and the similarities in ignition threshold between the single and double heater cases suggest that in the prescience of wind fluid dynamic effects are more influential than thermal effects.
        \begin{figure}
            \centering
             \begin{subfigure}[b]{0.49\textwidth}
                \centering
                \includegraphics[width=\textwidth, trim={27cm 0 0 0},clip]{Figures/recirculationComparison.png}
                \caption{0.5\si{\meter\per\second}}
                \label{subfig:comparisonLowWind}
            \end{subfigure}
            \begin{subfigure}[b]{0.49\textwidth}
                \centering
                \includegraphics[width=\textwidth, trim={0 0 27cm 0},clip]{Figures/recirculationComparison.png}
                \caption{5.8\si{\meter\per\second}}
                \label{subfig:comparisonHighWind}
            \end{subfigure}
            \caption{Comparison of streamlines for two different wind speeds with heaters spaced one diameter apart. Air flows from left to right. Images correspond to the bottom images in the left quadrants of Figure~\ref{fig:multiHeaterCFD}}
            \label{fig:streamlineComparison}
        \end{figure}
    Considering the cases where the wind speed was 5.8\si{\meter\per\second} but the upstream heater was not heated and is effectively an inert flow obstruction (Figure~\ref{fig:multiHeaterCFD} second row from the bottom). Following the trends of the previous case where both heaters were heated ignition would be expected to occur in the recirculation zone upstream of the actively heated heater. However, in these configurations ignition occurred on the downstream side of the upstream heater as indicated in Figure~\ref{fig:multiHeaterCFD} (second row from the bottom) and as shown in Figure~\ref{fig:unheatedUpstreamIgnition}. Figure~\ref{fig:unheatedUpstreamIgnition} shows images of the fuel beds shortly after ignition occurred for both the one diameter and five diameter spaced cases. 
        \begin{figure}
            \centering
             \begin{subfigure}[b]{0.49\textwidth}
                \centering
                \includegraphics[width=\textwidth]{Figures/1DupstreamInert.png}
                \caption{1D Spacing}
                \label{subfig:1DunheatedImage}
            \end{subfigure}
            \begin{subfigure}[b]{0.49\textwidth}
                \centering
                \includegraphics[width=\textwidth]{Figures/5DupstreamInert.png}
                \caption{5D Spacing}
                \label{subfig:5DunheatedImage}
            \end{subfigure}
            \caption{Images of fuel beds shortly after ignition near the upstream (left) heater. In both cases the wind speed was 5.8\si{\meter\per\second} and the upstream heater is unheated/inert. Air is travelling from left to right in both images.}
            \label{fig:unheatedUpstreamIgnition}
        \end{figure}
    In both heater cases ignition was occurs on the downstream side of the upstream heater after the pyrolysis front of the fuel bed propagated upstream and reached the inert heater. Effectively by not heating the upstream heater the location of ignition shifted from the upstream side of the upstream heater to the downstream side of the upstream heater. This is perhaps counter intuitive since ignition occurred in recirculation zones when both heaters were heated but the ignition locations for the inert upstream heater do not coincide with recirculation zones in the computational results shown in Figure~\ref{fig:multiHeaterCFD} (second row from bottom). In addition to ignition in an area that lacks recirculation in these cases ignition is also occurring further from the heat source than in the other heater configurations and wind speeds. Ignition far from the heat source, especially in the five diameter spaced case, suggests that the fuel bed is undergoing a smoldering to flaming transition. Ignition away from the heat source was not observed in the single heater cases presented in Chapter~\ref{part:manuscript2} thus, the upstream heater appears to be facilitating this transition. The smoldering to flaming transition near the inert heater is attributed to changes in the local fluid flow field as the smoldering front approaches the heater. Studies of the smoldering to flaming transition have found that the mass flux rate of smoldering fuel bed is sensitive to wind speed~\cite{Ohlemiller1990} and that the accumulation of pyrolyzates increases the likelihood of flaming~\cite{Stoliarov}. As the smoldering front approaches the upstream heater both a change in wind speed and an accumulation of pyrolyzates may occur creating a perturbation that ultimately results in ignition. Another potential contributor to ignition is the heater acting as an overhang which may contribute to the smoldering to flaming transition. Unfortunately, little information is available regarding the influence of overhangs on the smoldering to flaming transition but in this case overhangs seem to promote ignition~\cite{Santoso2019}. Despite the different ignition locations between the heater configurations with a wind speed of 5.8\si{\meter\per\second} the temperature required for a 50\% ignition probability differs by less than 10\% between all four configurations and in comparison to the single heater results from Chapter~\ref{part:manuscript2}. Thus, for the configurations tested, at a wind speed 5.8\si{\meter\per\second} has a more significant effect than the number, spacing, and whether or not both heaters are hot.
